---
title: "Statistical Learning for Data Science Exam"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error=TRUE)
```

The duration of the exam is 120 minutes. For most exercises you are required to provide R code: write your code in the assigned R code chunk. For some exercises you are required to provide a textual answer: type your answer in the space that follows 

**Your answer**.

In the following R chunk please fill in the required information otherwise the file will not be knitted

```{r, error=FALSE}
matricola <- 
#write your name within quotes, i.e. name <- "your name"
name <- "n"
#write your suname within quotes, i.e. name <- "your surname"
surname <- "c"
if (!exists("matricola") | !exists("name") | !exists("surname"))
  stop("Please, insert your information in the first chunk.")
```

```{r, cache=FALSE}
library(SLDS2425)
library(tidyverse)
data("beer")
```

# Exercise 1
A medium scale brewery has conducted a customer survey as a part of their market research strategy. A sample of $ 248$ beer drinkers have been selected and they were asked to taste 3 types of beer: light, dark and normal. Their preference was recorded as well as their demographics, including age and gender. Data are available in `beer` dataset. 


## a)
What is the mean age of survey participants? What is the mean age of female and male survey participants?

**Your answer**:
```{r}

```


## b) 
What is the overall distribution of beer preference among survey participants? 

**Your answer**:
```{r}

```

## c)
Use an appropriate numerical data summary and a graphical display to evaluate whether there is an association between age and the beer preference. 

**Your answer**:
```{r}


```

## d) 
What is the distribution of gender in the survey? 

**Your answer**:
```{r}


```


## e) 
Is there an association between gender and beer preference?  What data summaries and plots can you use to investigate and tentatively answer question? And if you would want to make a formal conclusion, what statistical test would you use? You do not need to perform the statistical test itself, just elaborate and justify your choice. 

**Your answer**:

```{r}

  
```


# Exercise 2
A data analyst has processed some data regarding supermarket sales (in dollars) of two categories of pasta, "FreshCutPasta" and "FreshNoodles". The data was collected during a time period of 10 days, precisely from 12/10/2024 to 12/19/2024 (the format of the date is month/day/year). In the chunk below you are given some information regarding the output of the analysis, in particular

  * the summary of a linear regression model estimated by the analyst. Some entries are intentionally set to "?" 
  * the vector of residuals (the variable containing the residuals is called `Resid`)
  * the vector with the values of the explanatory variable (the variable containing the values is called `pasta`)

To visualize the summary and get access to the variables containing the values of the residuals and of the explanatory variable *you have to compile the code*.
  
```{r echo=TRUE}
library(SLDS2425)
set.seed(matricola)
obj <- slds.fun(matricola, 1)
Resid <- obj$residuals
pasta <- obj$pasta
```

## (a) 
By using the information included in the summary above, say whether or not the explanatory variable is useful in explaining the response variable from a statistical perspective. Specifically, discuss the test statistic you use, the null and alternative hypothesis, and provide the conclusion (no coding needed, just plain text)

**Your answer**:



## (b)
By using the information included in the summary of the estimated linear model, provide the interpretation of the coefficients in the model. Motivate your answer (no coding needed, just plain text)

**Your answer**:

## (c)
By using all the available information, estimate the model displayed in the summary. Check whether there are influential observations.

**Your answer**:
```{r echo=TRUE}




```

## (d)
Use the model estimated in the previous point to predict the sales of the two categories of pasta on 12/24/2024. Motivate your answer (just text or text and code).

**Your answer**:



# Exercise 3
"Fiber Bits" is an internet service provider company. The company is concerned about the high attrition rate in the customer base. They want to get an idea on what are the main factors that lead to customer attrition.  

The company has collected around 10000 customer historical data from last three years. We need to build a model that identifies the customers who are most likely to leave, i.e. we need to quantify the chance of attrition for each of the customer. The model will be used on the active customers. To reduce the attrition rate, they have introduced vouchers and other benefits program. The objective is to identify the customers who are most likely to quit in next 2 years and try to retain them by offering free vouchers and benefits. 

Load the data by using the command `data("fiberbits")`. Here a short legend of the variables.

- `active_cust` Active - 1  (Customer Attrition=No)
Not Active – 0 (Customer Attrition=Yes)

- `income`  Estimated monthly income

- `months_on_network` Months on network (Months from the service start day)

- `Num_complaints` Total complaints till now

- `number_plan_changes` Number of times the service plan is changed

- `monthly_bill`  Average monthly bill

- `technical_issues_per_month` Number of technical issues per month

- `Speed_test_result` Percent of (Actual speed/Promised speed)


## (a) 
Run the chunk below (do not modify it). The outputs are contingency tables with the variable `active_cust` on the rows and `Num_complaints` on the columns. Which table would you use to assess whether the variable `Num_complaints` is worth to be included in a regression model with `active_cust` as response variable? Motivate you answer (just plain text)

```{r echo=TRUE}
library(SLDS2425)
data(fiberbits)
slds.fun(matricola, 2, fiberbits)
```


**Your answer**:

## (b) 
Run the chunk below (do not modify it). The output is a confusion matrix (predicted values on the rows, actual values on the columns) obtained from a classifier that was fit to the data. The matrix has been built by using data not used to estimate the model. By looking at the matrix argue about the ability of the classifier to correctly detect customers that quit. Motivate your answer (only text or text and code).

```{r echo=TRUE}
library(SLDS2425)
data(fiberbits)
slds.fun(matricola, 3, fiberbits)
```

**Your answer**:
```{r echo=TRUE}

```

## (c)
Split the data set into a  traning and test data sets. Estimate a statistical model for **the chance of a customer leaving** by using shrinkage methods. Select the best model according to a suitable criteria. 

**Your answer**:
```{r echo=TRUE}

```

## (d)
Build a classifier to identify the customers who are most likely to quit. Compare the performance of the classifier, using suitable metrics, with the classifier considered in point (b). Comment on the results.

**Your answer**:
```{r echo=TRUE}

```


# Session Info

Do not modify the chunk below
```{r}
Sys.info()
#last compiled
date()
#student info
paste(name, surname, "-", matricola, sep=" ")
```
